Vragenlijst thesis evaluatie

### 2015-02-27 Thesisfollowup/ThesisFollowUpSophieMarien.tex
Game formal definition

\section{Geef een uitgewerkt werkplan voor je thesis-activiteiten tussen nu en half juni (dan moet je je masterproef indienen). Dit werkplan moet PER TWEE WEKEN aangeven wat je precies gaat doen voor je masterproef. Maak het heel CONCREET: in het geval van implementatie, implementatie van welk deel precies, in het geval van het schrijven van tekst, het schrijven van welk hoofdstuk precies, e.d. Stel dit werkplan zelfstandig op, en bespreek het daarna met je begeleider, voordat je op deze mail antwoordt.}

Uitgewerkt plan steekt bij in de bijlage. In de timesheet file staat naast de werkuren aan wat ik ga werken.

\section{Schrijf een stuk tekst van minstens 5 pagina's waarin je de probleemstelling van je masterproef formuleert, gevolgd door de doelstellingen. Je verwerkt hier onder meer elementen van je literatuurstudie in.}

Ik heb afgesproken dat ik het stuk tekst tegen dit weekend uitschrijf. Hier is al een kleine inleiding.

Virussen zijn een groot probleem voor bedrijfsnetwerken. Ze kunnen gevoelige informatie verzamelen of een bedrijfsnetwerk platleggen. Gegeven de grote kost gebonden aan schade door malware, is het vinden van de juiste verdegingingsstrategie belangrijk. Het aanvallen en verdedigen van een bedrijfsnetwerk kan gezien worden als een spel, waarbij de verdediger en de aanvaller elk proberen de beste strategie te vinden.  Ik ga dus toelichten hoe het spel van aanvallen en verdedigen kan gedefinieerd worden als een variatie op het spel FlipIt. Dit laat toe om te onderzoeken wat de verschillende strategieën zijn van de netwerkbeheerder enerzijds en van de aanvaller die virussen zendt, anderzijds. Bedoeling is om in een volgende stap het spel verder te analyseren met behulp van Game Theory om te bepalen welke de dominerende strategieën zijn en of er zich Nash equilibria voordoen.

\section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}

Ik heb voldoende tijd in mijn thesis gestoken. Ik heb de timesheet aangepast naar mijn planning. Ik heb weinig examens en ben een week eerder klaar dan de lesvrije week dus dan kan ik nog verder werken aan mijn thesis. Voor deze komende weken heb ik minder uren ingepland omdat ik van andere vakken projecten moet afwerken en presentaties moet geven. De laatste week heb ik minder kunnen werken dan dat ik graag had willen doen omwille van een kaakgewrichtsontsteking die momenteel nog steeds niet over is. De volgende weken ga ik proberen meer te werken aan mijn thesis dan dat ik gepland heb, om mijn uren in te halen.

\section{Evalueer de vorderingen van je thesis t.o.v. het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}

Ik sta op schema volgens het schema dat ik zelf gemaakt heb. Ik heb mijn schema voor de komende weken aangepast.

\section{Stuur de ingevulde timesheet mee als bijlage}
\end{document}

>>>>

\section{Vergelijk wat je gedaan hebt de vorige weken met het opgestelde werkplan. Hoever sta je?}
Het waren examens vorige weken en er was een lesvrije week, dus veel vooruitgang heb ik niet geboekt in vergelijking met andere weken. Ik zat ook een beetje vast op een gegeven moment maar dan heb ik verder gewerkt aan de literatuurstudie en al beginnen te schrijven aan men thesis.
\section{Verfijn je werkplan (pas het eventueel aan) zodat dit werkplan PER WEEK, heel CONCREET opgeeft wat je van plan bent te doen.}
Dit werkplan heb ik concreter gemaakt in mijn timesheet. Dus daar kan het concretere werkplan teruggevonden worden
\section{Geef een outline van je thesistekst (Titels van hoofdstukken met per hoofdstuk 2 tot 5 regels over de inhoud). Dit mag je natuurlijk opstellen in overleg met je thesisbegeleider.}
\begin{description}
\item ABSTRACT
\item INTRODUCTION\\
\begin{itemize}
\item motivation game
\item contributions and results
\item lessons derived
\end{itemize}
\item CYBER SECURITY AND GAME THEORY
\item PROBLEM DEFINITION
\item MY GAME
\begin{itemize}
\item flipIt 
\item game motivation
\item examples
\item formal definitions
\item Default game is FlipIt
\item Multiple resources nash?
\item write out gain for different graphs and parameters
\end{itemize}
\item CONCLUDING REMARKS
\item GENERAL VARIATIONS AND EXTENSIONS
\item APPENDIX
\item REFERENCES
\end{description}
\section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
Voldoende. Ik ben maar 5 uur achtergeraakt op het schema in de week voor de lesvrije week maar toen ben ik ook ziek geweest.

\section{Evalueer de vorderingen van je thesis t.o.v. het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
Beetje achter op het schema, maar die achterstand kan snel ingehaald worden.
\section{Stuur de ingevulde timesheet mee als bijlage}

\end{document}

### 2015-02-27 doc/template/chap-2.tex
Game formal definition

\section{Figures}
\begin{figure}[hbtp]
\centering
\includegraphics[scale=1]{Images/FlipItDefault.jpg}
\caption{The FlipIt game where both players are playing periodically}
\label{fig:FLipItDefault}
\end{figure}
\todo{verwijzen naar de figuur \ref{fig:FLipItDefault}}

>>>>

\section{Figures}
\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.25]{Images/FlipItDefault.jpg}
\caption{The FlipIt game where both players are playing periodically}
\label{fig:FLipItDefault}
\end{figure}
\todo{verwijzen naar de figuur \ref{fig:FLipItDefault}}

\subsection{Strategies}

 In this subsection we go through the strategies used in FlipIt and the most important results. 
 \begin{table}
 \centering
 \begin{tabular}{ l | c  }
  \textbf{Categories} & \textbf{Classes of Strategies} \\
  \hline Non-adaptive (NA) & Exponential \\
  & Periodic \\
  & Renewal \\
  & General non-adaptive \\
  \hline Adaptive (AD) & Last move (LM) \\
  & Full History (FH) \\  
\end{tabular}
 \caption{Classes of strategies in FlipIt}
 \label{table:Strategies}
 \end{table}

There are two different kinds of strategies, the \textit{non-adaptive strategies} and the \textit{renewal strategies}. If there is no need for feedback for both of the players, we say that we have a non-adaptive strategy. Because the player does not receive any feedback during the game it will play in the same manner against every opponent. They are not dependent on the opponents movements. This means that they can already generate the time sequence for all the moves in advance.  But they can depend on some randomness because the non-adaptive strategies can be randomised. 
In this paper we will focus in the beginning on the non-adaptive strategies. Reasons behind this that a player (defender or attacker) rarely knows what the strategies are of his opponent. [If the attacker wants to move stealthily, it might have limited attack options FLIPTHEM]. \todo{nog redenen zoeken}\\
A renewal strategy is a non-adaptive strategy where the time intervals between two consecutive moves are generated by a renewal process. \\

 \begin{description}
 \item Periodic
 \item Non-Arithmetic Renewal
 \item Exponential
 \end{description}

\section{Formal definition Game}

In this section we provide the formal definition of the game and the notation that we will use throughout the paper.

\begin{description}
\item \textit{Players}  There are two players in the game, one is the defender and the other one is the attacker. They are respectively identified by 0 and 1.
\item \textit{Time}  The game starts at $t=0$ and continuous indefinitely as $t \rightarrow \infty$. The game is a continuous game.
\item \textit{Game State} There is also a time-dependent variable that represents the state of the game. $C=C(t)$ is either 0 if the Game is under control by the defender and 1 if the Game is under control by the attacker. We can also define the state of each resource by $C^{A}$ and $C^{D}$. If $C^{A} = 1$ then this means that the attacker has control over the resource, and 0 otherwise. For $C^{D}$ it is visa versa, $C^{D} = 1 - C^{A}$.
\item \textit{Graph} We represent the company network as a Graph $G = < V,E>$. G is an ordered pair where V denotes the set of resources or nodes in the network and E denotes the set of connections or links, which are a two-element subset of V. We use the notations resources and nodes interleaving in this paper.\\
We have N resources in the network. $N \in $  \todo{aanvullen}. This means we can denote the resources by:
\begin{center}
$V \in {V_{0}, V_{1}, V_{2}, ... , V_{N} }$
\end{center}
The set E of connections indicates if there is a link between two resources. We see the links as bidirectional so the total graph is undirected. If there is a link between resource $V_{n}$ and $V_{n+1}$ then there is also a link between $V_{n+1}$ and $V_{n}$.

\item \textit{Moves} Both players can make a move in the game. Moves done in a finite numbers of time in any finite time interval. Both players can play at any time they want, they can also play at the same time. If this happens the one that has control over the resource will keep having control over the resource.
This makes the game fully symmetric \todo{beter uitleggen}. The sequence of move times are denoted by the following infinite sequence:
\begin{center}
$t=t_{1},t_{2},t_{3},..$
\end{center}
Two move times can be the same because we allow players to move at the same time.
We can also denote the infinite sequence of times when player \textit{i}  moves. We write this as :
\begin{center}
$t=t_{i,1},t_{i,2},t_{i,3},..$ with \textit{i} $ \in $ $\lbrace 0,1 \rbrace$
\end{center}
The sequences $t_{1}$ and $t_{0}$ are disjoint subsets of the sequent t. 
We can also denote who made the \textit{k}th move by defining a sequence \textit{p} that denotes the sequence of who played:
\begin{center}
$p=p_{1},p_{2},p_{3}, .. $ with $p_{k}$ $\in$ $\lbrace 0,1 \rbrace$
\end{center}

\item \textit{Number of moves}  $n_{i}(t$ denotes the number of moves made by player \textit{i} up to and including time t. This means that 
\begin{center}
$n(t)=n_{1}(t) + n_{0}(t)$
\end{center}
is the sum of the number of moves made by the defender and the attacker up to and including time t.

\item \textit{Average move rate} We denote $\alpha_{i}(t)$ as the average move rate by player i:
\begin{center}
$\alpha_{i}(t) = n_{i}(t)/t$ with $t > 0$ and \textit{i} $ \in $ $\lbrace 0,1 \rbrace$
\end{center}

\item \textit{Period} We can also define the period in terms of the average move rate:
\begin{center}
$\delta_{i}=1/\alpha_{i}$
\end{center}

\item \textit{Cost} The cost is an important property of the game. In FlipIt for every player the cost of a move is denoted by $k_{i}$. These costs can be very different for every player. In this game we denote the players flipping cost for resource $V_{N}$ by $c_{i}^{V_{N}}$. \\
For the defender the cost will be either the cost of flipping every resource or the cost of flipping a subgroup of the resources.\\
For the attacker the cost will be the cost of dropping a virus on a node. The spreading of the virus will not imply an extra cost.

\item \textit{Utility} In FlipIt the Gain definition is the utility function. The Gain denotes the total time a player i has gained control over a resource. \todo{nu gain van een resource, moet voor verschillende resources zijn}
The Gain $G_{i}$ denotes players i total gain of a game, wich is the total time the player has gained control over a subset of resources thus controlling the game. If we sum up the total Gain of the attacker and the defender we end up with the time:
\begin{center}
$G_{1}(t) + G_{0}(t) = t$
\end{center}
\item \textit{Average gain rate} The average gain rate for player i is defined as
\begin{center}
$\gamma_{i}(t)= G_{i}(t)/t$
\end{center}

\subsection{Our Game parameters}
\begin{description}
\item \textit{Graph Matrix} We represent the graph of the network through a matrix $ A = |V| \times |V|$. The (i,j)-entry of the matrix A will have a 1 if there is a connection between node $V_{i}$ and node $V_{j}$. If we are working with an undirected graph the matrix will be symmetric. 
\item \textit{Attack Vector} We denote $X = 1 \times |V|$ as the attack vector.  
\item \textit{Reset vector} The reset vector will make sure that the right entries in the matrix become zero.
\item \textit{Cummulative Matrix} 
\item \textit{State Matrix} The State matrix $T(t) = |V| \times |V| $ will keep at every time t the state of the game.
\end{description}

\end{description}
